### 2022-06-02 chapters/interpretability/imli/introduction.tex
all papers added
minor edits in many parts of the thesis

\section{Introduction}

@@

In this paper, we study decision functions expressible in \emph{propositional logic}. In particular, we focus on learning logical formulas from data as classifiers and define interpretability in terms of the number of Boolean predicates that the formula contains. In the following, we illustrate an interpretable classifier that decides if a tumor cell is malignant or benign based on different features of tumors.

%\begin{subequations}
%	\begin{empheq}[box=\widefbox]{align*}
%	& \text{A tumor is malignant if}\\
%	&[(\text{compactness SE} < 0.1) \text{\textbf{ OR }}  \neg (0.1 \le \text{concave points} < 0.2)]
%	\text{\textbf{ \blue{AND} }} \\
%	&[\neg (0.2 \le \text{area} < 0.3) \text{\textbf{ OR }} \neg (0.1 \le \text{largest symmetry }< 0.2)]
%	\text{\textbf{ \blue{AND} }} \\
%	&[(0.6 \le \text{compactness} < 0.7) \text{\textbf{ OR }} (0.4 \le \text{concave points} < 0.5) \text{\textbf{ OR }}  (\text{area SE} \ge 0.1)\\
%	& \text{\textbf{ OR }} (0.3 \le \text{largest concavity} < 0.4) \text{\textbf{ OR }} (0.2 \le \text{largest area} < 0.3)]
%	\end{empheq}
%\end{subequations}

\boxed{
\begin{array}{rl}
	& \text{A tumor is malignant if}\\
	&[(\text{compactness SE} < 0.1) \text{\textbf{ OR }}  \neg (0.1 \le \text{concave points} < 0.2)]
	\text{\textbf{ \blue{AND} }} \\
	&[\neg (0.2 \le \text{area} < 0.3) \text{\textbf{ OR }} \neg (0.1 \le \text{largest symmetry }< 0.2)]
	\text{\textbf{ \blue{AND} }} \\
	&[(0.6 \le \text{compactness} < 0.7) \text{\textbf{ OR }} (0.4 \le \text{concave points} < 0.5) \text{\textbf{ OR }}  (\text{area SE} \ge 0.1)\\
	& \text{\textbf{ OR }} (0.3 \le \text{largest concavity} < 0.4) \text{\textbf{ OR }} (0.2 \le \text{largest area} < 0.3)]
\end{array}
}
\begin{example}
	We learn an interpretable classification rule in conjunctive normal form for predicting the classification of a tumor cell. We consider WDBC (Wisconsin Diagnostic Breast Cancer) dataset~\cite{agarap2018breast} to learn a CNF formula with three \emph{clauses} and nine Boolean \emph{predicates}. In the formula, clauses are connected by Boolean `AND', where each clause contains \emph{predicates} connected by Boolean `OR'. Informally, a tumor cell is malignant if at least one predicate in each clause becomes true.
\end{example}

@@

The MaxSAT problem is an optimization analog to the satisfiability (SAT) problem, which is complete for the class $ \mathrm{FP}^{\mathrm{NP}} $. Although the MaxSAT problem is NP-hard, dramatic progress has been made in designing solvers that can handle large-scale problems arising in practice. This has encouraged researchers to reduce several optimization problems into MaxSAT such as  optimal planning~\cite{robinson2010partial}, automotive configuration~\cite{walter2013applications}, data analysis in machine learning~\cite{berg2019applications}, and automatic test pattern generation for cancer therapy~\cite{lin2012application}.

The primary contribution of this paper is a MaxSAT-based formulation, called {\imli}, for learning interpretable classification rules expressible in propositional logic. For the simplicity of exposition, our initial focus is on learning classifiers expressible in CNF (Conjunctive Normal Form) formulas; we later discuss how the CNF learning formulation can be extended to popular interpretable classifiers such as decision lists and decision sets. In this paper, {\imli} provides precise control to jointly optimize the accuracy and the size of classification rules. {\imli} constructs and solves a MaxSAT query of $ \mathcal{O}(kn) $  size (number of clauses) to learn a $ k $-clause CNF formula on a dataset containing $ n $ samples. Naturally, the na\"ive formulation cannot scale to large values of $ n $ and $ k $. To scale {\imli} to large datasets, we propose an incremental learning technique along with the MaxSAT formulation. Our incremental learning is an integration of mini-batch learning and iterative rule-learning, which are studied separately in classical learning problems. In the presented incremental approach, {\imli} learns a $ k $-clause CNF formula using an iterative separate-and-conquer algorithm, where in each iteration, a single clause is learned by covering a part of the training data. Furthermore, to efficiently learn a single clause in a CNF, {\imli} relies on mini-batch learning, where it solves a sequence of smaller MaxSAT queries corresponding to each mini-batch.

@@

The presented framework {\imli} extends our prior papers ~\cite{ghosh19imli,MM18} in the following aspects. \cite{MM18} presented a MaxSAT-based formulation for interpretable classification rule-learning, which jointly optimizes between the accuracy and size of rules.  \cite{ghosh19imli} improved the MaxSAT formulation by unifying mini-batch learning. This unification allows learning on moderate-size datasets (about $ 50 $ thousand samples). This paper takes a significant step in terms of scalability by integrating both iterative learning and mini-batch learning inside the incremental technique. The result is a practical scalable learning framework that can classify datasets with $ 1 $ million samples and can potentially scale beyond that.

We organize the rest of the paper as follows. We discuss related literature in Section~\ref{interpretability_imli_sec:related} and preliminaries in Section~\ref{interpretability_imli_sec:preliminaries}. In Section~\ref{interpretability_imli_sec:problem}, we formally define our learning problem and provide a MaxSAT-based solution in Section~\ref{interpretability_imli_sec:baseline}. We improve the MaxSAT-based formulation through an incremental learning technique in Section~\ref{interpretability_imli_sec:incremental_learning}. In Section~\ref{interpretability_imli_sec:application}, we apply {\imli} in learning different interpretable classifiers. We then discuss our experimental results in Section~\ref{interpretability_imli_sec:experiments} and conclude in Section~\ref{interpretability_imli_sec:conclusion}.

>>>>

%\section{Introduction}

@@

In this chapter, we study decision functions expressible in \emph{propositional logic}. In particular, we focus on learning logical formulas from data as classifiers and define interpretability in terms of the number of Boolean predicates that the formula contains. In the following, we illustrate an interpretable classifier that decides if a tumor cell is malignant or benign based on different features of tumors.

\boxed{
\begin{array}{rl}
	& \text{A tumor is malignant if}\\
	&[(\text{compactness SE} < 0.1) \text{\textbf{ OR }}  \neg (0.1 \le \text{concave points} < 0.2)]
	\text{\textbf{ \blue{AND} }} \\
	&[\neg (0.2 \le \text{area} < 0.3) \text{\textbf{ OR }} \neg (0.1 \le \text{largest symmetry }< 0.2)]
	\text{\textbf{ \blue{AND} }} \\
	&[(0.6 \le \text{compactness} < 0.7) \text{\textbf{ OR }} (0.4 \le \text{concave points} < 0.5) \text{\textbf{ OR }}  (\text{area SE} \ge 0.1)\\
	& \text{\textbf{ OR }} (0.3 \le \text{largest concavity} < 0.4) \text{\textbf{ OR }} (0.2 \le \text{largest area} < 0.3)]
\end{array}
}

\begin{example}
	\normalfont
	We learn an interpretable classification rule in conjunctive normal form for predicting the classification of a tumor cell. We consider WDBC (Wisconsin Diagnostic Breast Cancer) dataset~\cite{agarap2018breast} to learn a CNF formula with three \emph{clauses} and nine Boolean \emph{predicates}. In the formula, clauses are connected by Boolean `AND', where each clause contains \emph{predicates} connected by Boolean `OR'. Informally, a tumor cell is malignant if at least one predicate in each clause becomes true.
\end{example}

@@

The MaxSAT problem is an optimization analog to the satisfiability (SAT) problem, which is complete for the class $ \mathrm{FP}^{\mathrm{NP}} $. Although the MaxSAT problem is NP-hard, dramatic progress has been made in designing solvers that can handle large-scale problems arising in practice. This has encouraged researchers to reduce several optimization problems into MaxSAT such as  optimal planning~\cite{robinson2010partial}, automotive configuration~\cite{walter2013applications}, group-testing~\cite{CGSM2020}, data analysis in machine learning~\cite{berg2019applications}, and automatic test pattern generation for cancer therapy~\cite{lin2012application}.

The primary contribution of this chapter is a MaxSAT-based formulation, called {\imli} (\textbf{I}ncremental \textbf{M}axSAT-based \textbf{L}earning of \textbf{I}nterpretable classification rules), for learning interpretable classification rules expressible in propositional logic. For the simplicity of exposition, our initial focus is on learning classifiers expressible in CNF (Conjunctive Normal Form) formulas; we later discuss how the CNF learning formulation can be extended to popular interpretable classifiers such as decision lists and decision sets. In this chapter, {\imli} provides precise control to jointly optimize the accuracy and the size of classification rules. {\imli} constructs and solves a MaxSAT query of $ \mathcal{O}(kn) $  size (number of clauses) to learn a $ k $-clause CNF formula on a dataset containing $ n $ samples. Naturally, the na\"ive formulation cannot scale to large values of $ n $ and $ k $. To scale {\imli} to large datasets, we propose an incremental learning technique along with the MaxSAT formulation. Our incremental learning is an integration of mini-batch learning and iterative rule-learning, which are studied separately in classical learning problems. In the presented incremental approach, {\imli} learns a $ k $-clause CNF formula using an iterative separate-and-conquer algorithm, where in each iteration, a single clause is learned by covering a part of the training data. Furthermore, to efficiently learn a single clause in a CNF, {\imli} relies on mini-batch learning, where it solves a sequence of smaller MaxSAT queries corresponding to each mini-batch.

@@

%The presented framework {\imli} extends our prior chapters ~\cite{ghosh19imli,MM18} in the following aspects. \cite{MM18} presented a MaxSAT-based formulation for interpretable classification rule-learning, which jointly optimizes between the accuracy and size of rules.  \cite{ghosh19imli} improved the MaxSAT formulation by unifying mini-batch learning. This unification allows learning on moderate-size datasets (about $ 50 $ thousand samples). This chapter takes a significant step in terms of scalability by integrating both iterative learning and mini-batch learning inside the incremental technique. The result is a practical scalable learning framework that can classify datasets with $ 1 $ million samples and can potentially scale beyond that.

%We organize the rest of the chapter as follows. We discuss related literature in Section~\ref{interpretability_imli_sec:related} and preliminaries in Section~\ref{interpretability_imli_sec:preliminaries}. In Section~\ref{interpretability_imli_sec:problem}, we formally define our learning problem and provide a MaxSAT-based solution in Section~\ref{interpretability_imli_sec:baseline}. We improve the MaxSAT-based formulation through an incremental learning technique in Section~\ref{interpretability_imli_sec:incremental_learning}. In Section~\ref{interpretability_imli_sec:application}, we apply {\imli} in learning different interpretable classifiers. We then discuss our experimental results in Section~\ref{interpretability_imli_sec:experiments} and conclude in Section~\ref{interpretability_imli_sec:conclusion}.

### 2022-06-02 chapters/interpretability/imli/iterative_learning.tex
minor edits in many parts of the thesis

\subsection{Learning Decision Lists}
In {\imli}, we apply an iterative learning approach for efficiently learning a  decision list. A decision list  $ \Rule_L $ is a list of pairs $ (C_1, v_1), \dots, (C_k, v_k) $, where we propose to learn one pair in each iteration.  We note that the clause $ C_i $ is a conjunction of literals\textemdash equivalently, a single clause DNF formula. Hence, our task is to deploy {\imli} to efficiently learn a single clause DNF formula $ C_i $. In particular, we opt to learn this clause for the majority class, say $ v_i $, in the training dataset by setting the majority samples as class $ 1 $ and all other samples as class $ 0 $. As a result, even if the MaxSAT-based learning, presented in this paper, is targeted for binary classification, we can learn a multi-class decision list in {\imli}.

>>>>

\subsection{Learning Decision Lists}
In {\imli}, we apply an iterative learning approach for efficiently learning a  decision list. A decision list  $ \Rule_L $ is a list of pairs $ (C_1, v_1), \dots, (C_k, v_k) $, where we propose to learn one pair in each iteration.  We note that the clause $ C_i $ is a conjunction of literals\textemdash equivalently, a single clause DNF formula. Hence, our task is to deploy {\imli} to efficiently learn a single clause DNF formula $ C_i $. In particular, we opt to learn this clause for the majority class, say $ v_i $, in the training dataset by setting the majority samples as class $ 1 $ and all other samples as class $ 0 $. As a result, even if the MaxSAT-based learning, presented in this chapter, is targeted for binary classification, we can learn a multi-class decision list in {\imli}.

### 2022-06-02 chapters/interpretability/imli/problem_formulation.tex
minor edits in many parts of the thesis

 Our goal is to learn classifiers that balance two 
 goals: of  being accurate but also interpretable.  
 Various notions of interpretability have been proposed in the context of   classification problems. A common proxy for interpretability in the context of decision rules 
 is the sparsity of rule. For instance, a rule involving fewer literals is highly interpretable.  In this work, we minimize the total number of literals in all clauses, as discussed in Section~\ref{interpretability_imli_sec:preliminaries}, which motivates us to  find $ \Rule  $ with minimum  $ |\Rule| $. Let $ \Rule $ classify all samples correctly during training. Among all the classification rules that classify all samples correctly,  we choose the sparsest (most interpretable) such $ \Rule $.

>>>>

 Our goal is to learn classifiers that balance two 
 goals: of  being accurate but also interpretable.  
 Various notions of interpretability have been proposed in the context of   classification problems. A common proxy for interpretability in the context of decision rules 
 is the sparsity of rule. For instance, a rule involving fewer literals is highly interpretable.  In this work, we minimize the total number of literals in all clauses, which motivates us to  find $ \Rule  $ with minimum  $ |\Rule| $. Let $ \Rule $ classify all samples correctly during training. Among all the classification rules that classify all samples correctly,  we choose the sparsest (most interpretable) such $ \Rule $.

### 2022-06-02 chapters/interpretability/relaxed-cnf/conclusion.tex
all papers added
minor edits in many parts of the thesis

\section{Conclusion}
\label{interpretability_crr_sec:conclusion}

In this paper, we proposed an efficient combinatorial framework called {\crr}, for learning relaxed-CNF  classification rules, that are  more expressive  than CNF/DNF rules.  {\crr}  uses a novel integration of mini-batch learning procedure with the MILP framework to learn sparse relaxed-CNF rules. Our experimental results demonstrate that {\crr} is able to learn relaxed-CNF  rules with higher accuracy and more concise representation than CNF rules. Moreover, the generated rules are sparser than decision lists in large datasets.

>>>>

\section{Chapter Summary}
\label{interpretability_crr_sec:conclusion}

In this chapter, we proposed an efficient combinatorial framework called {\crr}, for learning relaxed-CNF  classification rules, that are  more expressive  than CNF/DNF rules.  {\crr}  uses a novel integration of mini-batch learning procedure with the MILP framework to learn sparse relaxed-CNF rules. Our experimental results demonstrate that {\crr} is able to learn relaxed-CNF  rules with higher accuracy and more concise representation than CNF rules. Moreover, the generated rules are sparser than decision lists in large datasets.

### 2022-06-02 chapters/interpretability/relaxed-cnf/experiment.tex
minor edits in many parts of the thesis

%%\vspace{-3pt}
\subsection{Experiment Methodology}
We perform experiments on a high-performance computer cluster, where each node consists of E$ 5$-$2690\text{ v}3 $ CPU with $ 24 $ cores, $ 96 $ GB of RAM. Each experiment is run on four cores of a node with $ 16 $ GB memory.   We compare the performance of  $ \crr $   with  state-of-the-art classifiers, e.g. $ \IMLI  $~\cite{ghosh19incremental}, RIPPER~\cite{C1995}, BRS~\cite{WRDLKM2017},  random forest (RF), support vector classifier (SVC), nearest neighbors classifiers ($ k $-NN), and $ l_1 $ penalized logistic regression (LR). Among them, {\IMLI}, BRS, and RIPPER are rule-based classifiers. In particular,
{\IMLI} generates classification rules in CNF using a MaxSAT-based formulation and we use Open-WBO \cite{MML2014} as the MaxSAT solver for {\IMLI}. We compare with propositional rule learning algorithm RIPPER, which is implemented in WEKA~\cite{HFHPRW2009} and generates classification rules in the form of decision lists. BRS is a Bayesian framework for generating rule sets expressed as DNF. For  other classifiers, we use the Scikit-learn module of Python~\cite{PVGMTGB2011}. For all  classifiers, we set the training cut-off time to $ 1500 $  {seconds}.

>>>>

%%\vspace{-3pt}
\subsection{Experiment Methodology}
We perform experiments on a high-performance computer cluster, where each node consists of E$ 5$-$2690\text{ v}3 $ CPU with $ 24 $ cores, $ 96 $ GB of RAM. Each experiment is run on four cores of a node with $ 16 $ GB memory.   We compare the performance of  $ \crr $   with  state-of-the-art classifiers, e.g. $ \IMLI  $~\cite{GM2019}, RIPPER~\cite{C1995}, BRS~\cite{WRDLKM2017},  random forest (RF), support vector classifier (SVC), nearest neighbors classifiers ($ k $-NN), and $ l_1 $ penalized logistic regression (LR). Among them, {\IMLI}, BRS, and RIPPER are rule-based classifiers. In particular,
{\IMLI} generates classification rules in CNF using a MaxSAT-based formulation and we use Open-WBO \cite{MML2014} as the MaxSAT solver for {\IMLI}. We compare with propositional rule learning algorithm RIPPER, which is implemented in WEKA~\cite{HFHPRW2009} and generates classification rules in the form of decision lists. BRS is a Bayesian framework for generating rule sets expressed as DNF. For  other classifiers, we use the Scikit-learn module of Python~\cite{PVGMTGB2011}. For all  classifiers, we set the training cut-off time to $ 1500 $  {seconds}.

### 2022-06-02 chapters/interpretability/relaxed-cnf/ilp_formulation.tex
minor edits in many parts of the thesis

After applying entropy-based discretization on continuous features,  the dataset  contains only categorical features, that can be converted to binary features using one-hot encoding as in \cite{ghosh19incremental,LKCL2019}. In this encoding, a Boolean vector is introduced with cardinality equal to the number of distinct categories. Let a categorical feature have three  categories `red',`green', and `yellow'. In one-hot encoding, samples with category-value `red',`green', and `yellow' would be converted into binary features while taking values $ 100 $, $ 010 $, and $ 001 $, respectively.

>>>>

After applying entropy-based discretization on continuous features,  the dataset  contains only categorical features, that can be converted to binary features using one-hot encoding as in \cite{LKCL2019}. In this encoding, a Boolean vector is introduced with cardinality equal to the number of distinct categories. Let a categorical feature have three  categories `red',`green', and `yellow'. In one-hot encoding, samples with category-value `red',`green', and `yellow' would be converted into binary features while taking values $ 100 $, $ 010 $, and $ 001 $, respectively.

### 2022-06-02 chapters/interpretability/relaxed-cnf/introduction.tex
minor edits in many parts of the thesis

The widespread adoption of prediction systems in various safety-critical domains such as medical diagnosis, law, education, and many others has led to the increased importance of presenting the decision functions in interpretable representations~\cite{K2001,MVBB2005,SAD2015,S2014,TV2013}.  To enable  safe, robust, and trustworthy integration of such systems, the end users require them to support interpretability, privacy, and fairness in decision-making~\cite{DF2018,VML2012,WRLKM2015,ZUR2017}. In this context,  rule-based representations are considered interpretable in presenting the decision functions to users~\cite{LKCL2019,MM18,WR2015,WRDLKM2017}. A recent body of work has studied sparsity-inducing objectives for classification rules in CNF ({Conjunctive Normal Form}) or DNF ({Disjunctive Normal Form}) and demonstrated that they often achieve high interpretability\textemdash defined in terms of rule-sparsity\textemdash with a minimal sacrifice in classification accuracy ~\cite{ghosh19incremental,LKCL2019,MM18}. Although CNF/DNF rules are considered interpretable, they are less expressive compared to Boolean cardinality constraints in propositional logic. A Boolean cardinality constraint allows one to express numerical bounds on Boolean variables~\cite{sinz2005towards}. In this work, we introduce a novel formulation of interpretable classification rules, namely \emph{relaxed-CNF}, which achieve benefits of  both worlds: it is interpretable similar to CNF/DNF  but more expressible by allowing cardinality constraints in the representation.

@@

{Through a comprehensive experimental evaluation over datasets from the UCI and Kaggle repository, we observe that {\crr} with relaxed-CNF rules achieves an improved trade-off between accuracy and rule sparsity and scales to datasets with more than $ 10^5 $ samples. More significantly, {\crr} generates relaxed-CNF rules with higher accuracy than CNF rules generated by~\cite{ghosh19incremental}. Furthermore, compared to decision lists generated by~\cite{C1995}, relaxed-CNF rules are sparser in large datasets.}
%can also reach the same level of accuracy compared to plain CNF/DNF rules with smaller relaxed-CNF rules. 
%
%We have conducted a comprehensive experimental study on the datasets from the UCI repository. Our empirical study demonstrates  an increase in the prediction accuracy for choosing relaxed-CNF rules over regular CNF rules. The incremental approach achieves runtime scalability without loss of accuracy.

@@

%KSM: Please feel free to comment out the text below if it exceeds page limit
The rest of the chapter is organized as follows. We discuss related works in Section~\ref{interpretability_crr_sec:related_work}, introduce the notation and preliminaries in Section~\ref{interpretability_crr_sec:preliminaries} and formulate the problem in Section~\ref{interpretability_crr_sec:problem}. We describe the main contribution of this chapter, $ \crr $,  in Section~\ref{interpretability_crr_sec:ilp} and extend to an incremental mini-batch learning in Section~\ref{interpretability_crr_sec:inc_ilp}. We then describe the experimental results in Section ~\ref{interpretability_crr_sec:experiment} and conclude in Section~\ref{interpretability_crr_sec:conclusion}.

>>>>

The widespread adoption of prediction systems in various safety-critical domains such as medical diagnosis, law, education, and many others has led to the increased importance of presenting the decision functions in interpretable representations~\cite{K2001,MVBB2005,SAD2015,S2014,TV2013}.  To enable  safe, robust, and trustworthy integration of such systems, the end users require them to support interpretability, privacy, and fairness in decision-making~\cite{DF2018,VML2012,WRLKM2015,ZUR2017}. In this context,  rule-based representations are considered interpretable in presenting the decision functions to users~\cite{LKCL2019,WR2015,WRDLKM2017}. A recent body of work has studied sparsity-inducing objectives for classification rules in CNF ({Conjunctive Normal Form}) or DNF ({Disjunctive Normal Form}) and demonstrated that they often achieve high interpretability\textemdash defined in terms of rule-sparsity\textemdash with a minimal sacrifice in classification accuracy ~\cite{GM2019,LKCL2019,MM18}. Although CNF/DNF rules are considered interpretable, they are less expressive compared to Boolean cardinality constraints in propositional logic. A Boolean cardinality constraint allows one to express numerical bounds on Boolean variables~\cite{sinz2005towards}. In this work, we introduce a novel formulation of interpretable classification rules, namely \emph{relaxed-CNF}, which achieve benefits of  both worlds: it is interpretable similar to CNF/DNF  but more expressible by allowing cardinality constraints in the representation.

@@

{Through a comprehensive experimental evaluation over datasets from the UCI and Kaggle repository, we observe that {\crr} with relaxed-CNF rules achieves an improved trade-off between accuracy and rule sparsity and scales to datasets with more than $ 10^5 $ samples. More significantly, {\crr} generates relaxed-CNF rules with higher accuracy than CNF rules generated by~\cite{GM2019}. Furthermore, compared to decision lists generated by~\cite{C1995}, relaxed-CNF rules are sparser in large datasets.}
%can also reach the same level of accuracy compared to plain CNF/DNF rules with smaller relaxed-CNF rules. 
%
%We have conducted a comprehensive experimental study on the datasets from the UCI repository. Our empirical study demonstrates  an increase in the prediction accuracy for choosing relaxed-CNF rules over regular CNF rules. The incremental approach achieves runtime scalability without loss of accuracy.

@@

%KSM: Please feel free to comment out the text below if it exceeds page limit
%The rest of the chapter is organized as follows. We discuss related works in Section~\ref{interpretability_crr_sec:related_work}, introduce the notation and preliminaries in Section~\ref{interpretability_crr_sec:preliminaries} and formulate the problem in Section~\ref{interpretability_crr_sec:problem}. We describe the main contribution of this chapter, $ \crr $,  in Section~\ref{interpretability_crr_sec:ilp} and extend to an incremental mini-batch learning in Section~\ref{interpretability_crr_sec:inc_ilp}. We then describe the experimental results in Section ~\ref{interpretability_crr_sec:experiment} and conclude in Section~\ref{interpretability_crr_sec:conclusion}.

### 2022-06-02 chapters/interpretability_background.tex
all papers added
minor edits in many parts of the thesis

In this chapter, we discuss preliminaries for learning interpretable rule-based classifiers. Throughout this part of the thesis, we use capital  boldface letters such as $\mathbf{X}$ to denote matrices while lower boldface letters $\mathbf{x}$ are reserved for vectors/sets.

@@

Given a weighted-CNF formula $F \triangleq \bigvee_i C_i$ with weight $ \weight{C_i} $, the weighted-partial MaxSAT problems finds an optimal assignment $\sigma^*$ that achieves the minimum weight. Formally,  $\sigma^* = \MaxSAT(F,\weight{\cdot})$ if $\forall \sigma \neq \sigma^*, \weight{\sigma^*} \leq \weight{\sigma}$. The optimal weight of the MaxSAT problem $  \weight{\sigma^*} $ is infinity ($ \infty $) when at least one hard clause becomes unsatisfied.  Therefore,  the weighted-partial MaxSAT problem finds $ \sigma^* $ that satisfies all hard clauses and as many soft clauses as possible such that the total weight of unsatisfied soft clauses is minimum. In this paper, we reduce the classification problem  to the solution of a weighted-partial MaxSAT problem. The knowledge of the inner workings of {MaxSAT} solvers is  not required for this paper.

\subsection*{\red{Mixed Integer Linear Programming (MILP)}}
\red{[Not required]}The MILP problem is an optimization problem operating on both integer and real-valued variables. Given an objective function and a set of constraints, a (minimization) MILP finds an optimal assignment of variables that minimizes the objective function.

>>>>

In this chapter, we discuss preliminaries for learning interpretable rule-based classifiers. Throughout this part of the thesis, we use capital  boldface letters such as $\mathbf{X}$ to denote matrices while lower boldface letters such as $\mathbf{x}$ are reserved for vectors/sets.

@@

Given a weighted-CNF formula $F \triangleq \bigvee_i C_i$ with weight $ \weight{C_i} $, the weighted-partial MaxSAT problems finds an optimal assignment $\sigma^*$ that achieves the minimum weight. Formally,  $\sigma^* = \MaxSAT(F,\weight{\cdot})$ if $\forall \sigma \neq \sigma^*, \weight{\sigma^*} \leq \weight{\sigma}$. The optimal weight of the MaxSAT problem $  \weight{\sigma^*} $ is infinity ($ \infty $) when at least one hard clause becomes unsatisfied.  Therefore,  the weighted-partial MaxSAT problem finds $ \sigma^* $ that satisfies all hard clauses and as many soft clauses as possible such that the total weight of unsatisfied soft clauses is minimum. In this chapter, we reduce the classification problem  to the solution of a weighted-partial MaxSAT problem. The knowledge of the inner workings of {MaxSAT} solvers is  not required for this chapter.

%
%\subsection*{\red{Mixed Integer Linear Programming (MILP)}}
%\red{[Not required]}The MILP problem is an optimization problem operating on both integer and real-valued variables. Given an objective function and a set of constraints, a (minimization) MILP finds an optimal assignment of variables that minimizes the objective function.
